---
title: "Lab 1 Assignment"
author: "Dhananjay"
date: "2023-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1. Install R and R-Studio

Successfully installed R and R-Studio on my Windows 11 and tested the setup by generating a simple **hello world** rmd file.

## Q2. Get the Data:

* Inspecting **lab1.dat** in the repository
  - Yes, the column headers is included in the file.
  - Data columns are separated by ',' (comma)
  - No, there is no comments in the data

* Loading the data file into a data frame in R.

```{r}
dataFrame = read.table("lab1.dat", header=TRUE, sep=',', comment.char = "#")
```

```{r}
head(dataFrame)
```

## Q3. Peruse the Data:

* Print the name of the **variables** in the data table (hint: ?colnames)
```{r}
colnames(dataFrame)
```

* Provide a global **summary** of the data table (hint: ?summary)
```{r}
summary(dataFrame)
```

## Q4. Investigating the WineType variable:

* List all the values of the WineType variable in the data table
```{r}
print(dataFrame$WineType)
```

* Compute the levels of that factor (hint: ?levels)
```{r}
levels(factor(dataFrame$WineType))
```

* Compute the number of records of each level (hint: ?table)
```{r}
table(factor(dataFrame$WineType))
```

## Q5. Investigate the ColorIntensity variable:

* Compute mean of the ColorIntensity variable  (hint:  ?mean)
```{r}
mean(dataFrame$ColorIntensity)
```
  
* Compute variance of the ColorIntensity variable  (hint: ?var)
```{r}
var(dataFrame$ColorIntensity)
```

* Compute standard deviation of the ColorIntensity variable
```{r}
sd(dataFrame$ColorIntensity)
```

* Compute 95% confidence interval of the ColorIntensity variable (hint: ?t.test)
```{r}
n = dim(dataFrame)[1]                 # Sample Size
xb = mean(dataFrame$ColorIntensity)     # x_bar
s  = sqrt(var(dataFrame$ColorIntensity)) # standard deviation
z  = qnorm(0.975)                    # Critical value for 95%
xb - z*s/sqrt(n)
xb +z*s/sqrt(n)
```

```{r}
t.test(dataFrame$ColorIntensity)$conf.int
```
  
## Q6. Investigate the Hue variable for WineType Class1:

* Compute variance of the Hue variable across only the records with WineType=Class1 (hint: ?subset)
```{r}
hue_data = subset(dataFrame, WineType=="Class1", select = c(Hue))
print(hue_data)
var(hue_data)
```

* Compute standard deviation of the Hue variable across only the records with WineType=Class1
```{r}
print(hue_data)
hue_data_unlisted = unlist(hue_data)
print(hue_data_unlisted)
sd(hue_data_unlisted)
```

```{r}
hue_data_numeric = as.numeric(unlist(hue_data))
class(hue_data_numeric)
print(hue_data_numeric)
sd(hue_data_numeric)
```

* Compute 95% confidence interval of the Hue variable across only the records with WineType=Class1
```{r}
n = dim(dataFrame)[1]                 # Sample Size
xb = mean(hue_data_numeric)     # x_bar
s  = sqrt(var(hue_data_numeric)) # standard deviation
z  = qnorm(0.975)                    # Critical value for 95%
xb - z*s/sqrt(n)
xb +z*s/sqrt(n)
```

```{r}
t.test(hue_data_numeric)$conf.int
```
		
## Q7. Perform simple math operations:

* Subtract 120 from 536
```{r}
536-120
```

* Multiply 3 and 7
```{r}
3*7
```

* Divide 12 by 3
```{r}
12/3
```

* Take the base 2 logarithm of 128
```{r}
log(128,2)
```

* Compute nine cubed
```{r}
9^3
```
  
## Q8. Perform math operations on data columns:

* Compute the vector resulting from a pair-wise multiplication of all Hue variable values with ColorIntensity values
```{r}
hue_only_data = dataFrame$Hue
print(hue_only_data)

color_only_data = dataFrame$ColorIntensity
print(color_only_data)

pair_wise_mult_result = hue_only_data * color_only_data
print(pair_wise_mult_result)
```


* Sum the resulting vector from the previous step (hint:  ?sum)
```{r}
sum(pair_wise_mult_result)
```
  
## Q9. Drawing random numbers:

* Draw 10 numbers from the uniform distribution, between 0.0 and 10.0 (hint:  ?runif)
```{r}
runif(10, min = 0.0, max = 10.0)
```

* Draw 10 numbers from the Normal distribution, mean of 1.0, standard deviation of 0.3  (hint:  ?rnorm)
```{r}
rnorm(10, mean = 1.0, sd = 0.3)
```

* Compute PDF result for the standard Normal distribution at the value 0.12 (hint: ?dnorm)
```{r}
dnorm(0.12)
```

* Compute the probability that a value drawn from the standard
  - The question is confusing to me. I think that few parameters are missing. 
```{r}
qnorm(0.12)
```

* Normal distribution will be less than 1.96  (hint:  ?pnorm)
```{r}
value = 1.96
pnorm(value)
```
  
## Q10. Simulation:
* Perform a Monte Carlo sampling simulation of customers arriving at a queue over the period of 10 minutes, assuming this process is governed by the Poisson distribution and that 2 customers arrive per minute on average (hint:  ?rpois)
```{r}
rpois(10, lambda=2)
```

* Compute the total number of people standing in line at each minute for one such simulation?  (hint: ?cumsum)
```{r}
total = rpois(10, lambda=2)
cumsum(total)
```

* Enter the R code below, which defines a function.  Call that function using n=10, then again using n=10000, then again using n=10000000.  What does the function estimate?

```{r}
estimate <- function(n) {
  return (4*sum((runif(n)^2 + runif(n)^2) < 1)/n)
}
estimate(10)
estimate(10000)
estimate(10000000)
```

The function **estimate** returns the output as per the defined equation based on the input provided by the user.

